---
title: "ÄoB analysis of Butterflydata"
author: "Fabian Roger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
---


```{r}
library(ggplot2)
library(ggthemes)
library(sf)
library(dplyr)
library(tidyr)
library(lubridate)
library(RColorBrewer)
library(ggsn)
library(iNEXT)
library(readxl)
library(vegan)
library(here)
library(phyloseq)

#source("estimateD_func.r")
```

#import data

```{r}
load("RMO_clean_data.RData")

NILS <- read_sf("NILS_sw99_NILS_Stratum_SW99.shp")

Spec_cat <- read_excel("Species_meta_check.xlsx", sheet = "final")

```


```{r}
##CHECK WHY ÖREBRO IS FALLING OUT (LENGTH)

RMO %>% 
  dplyr::select(SiteID, SiteType, length, county, County_text, date) %>%
  distinct() %>% 
  group_by(SiteID, SiteType, county, County_text) %>% 
  mutate(date = 1:n()) %>% 
  pivot_wider(names_from = date, values_from = length) %>% 
  dplyr::select(1:6) %>%
  #arrange(SiteID)
  group_by(SiteType, county, County_text) %>% 
  summarise(sum(is.na(`1`)) / n()) 

RMO

library(sf)
library(here)
test <- read_sf(here("length", "LST_at_the_beginning.shp"))
test$MERGE_SRC %>% unique()


test <- read_sf(here("length", "FjarilsInv_ytor_tom2019.shp"))

```

#prepare data


## metadata
Prepare RMO data

split data into metadata (`RMO_meta`) and a Species matrix (`RMO_spec`), identified by unique ID

```{r}

RMO_meta <- 
  RMO %>%
  select(-SpeciesID, -SpeciesText, -count) %>% 
  distinct() %>% 
  mutate(ID = paste(date, SiteID, segment, sep = "_")) %>% 
  left_join(st_set_geometry(RMO_spat, NULL))  #join AOB site ID

```


we split the data into two rounds. 
`round_1` : 2009 and 2013
`round_2` : 2013 and 2019


```{r}
RMO_meta %>% 
  select(SiteID, date, county) %>% 
  mutate(county = ifelse(county == "SLU", "SLU", "Län")) %>% 
  mutate(date = year(date)) %>% 
  distinct() %>% 
  mutate(yes =1 ) %>% 
  pivot_wider(names_from = date, values_from = yes, values_fill = 0) %>% 
  select(-SiteID) %>% 
  group_by_all() %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(case = 1:n()) %>% 
  pivot_longer(cols = starts_with("20"),names_to = "date") %>% 
  mutate(date = as.numeric(date)) %>% 
  filter(value == 1) %>% 
  select(-value) %>% 
  distinct() %>% 
  ggplot(aes(x = date, y = case, fill = n, colour = n))+
  facet_wrap(~county, scales = "free_y") +
  geom_point(shape = 21)+
  geom_line(aes(group = case))+
  scale_fill_viridis_c()+
  scale_colour_viridis_c()+
  geom_vline(xintercept = 2013.5, linetype = "dashed", colour = "darkred")
```  

adding `round` variable
```{r}
RMO_meta <- 
RMO_meta %>% 
  mutate(round = case_when(year(date) <= 2013 ~ "round_1",
                           TRUE ~ "round_2"))
```


##sites with missing length data

How many sites have no length information at all?
```{r}

RMO_meta %>% 
  select(SiteText, SiteType, county, County_text, length) %>% 
  group_by(SiteText) %>% 
  summarise(SiteType = unique(SiteType),
            county = unique(county),
            County_text = unique(County_text),
            length = all(is.na(length))) %>% 
  group_by(county, SiteType) %>% 
  summarise(prct_no_len = round(sum(length)/n()*100)) %>% 
  left_join(filter(distinct(select(RMO_meta, county, County_text)), county != "SLU")) %>% 
  select(County_text, county, SiteType, prct_no_len)
  
```

N excluded site without any length information
```{r}
RMO_meta %>% 
  filter(!is.na(length)) %>% 
  pull(SiteText) %>% 
  unique() %>% 
  length() %>% 
  `-`(length(unique(RMO_meta$SiteText)), .)
```

I don't exclude sites without length info as we will base richness / diversity on abundance based coverage indices, not standardize by length

## species matrix
```{r}

RMO_spec <- 
  RMO %>%
  mutate(ID = paste(date, SiteID, segment, sep = "_")) %>% 
  select(ID, SpeciesText, count) %>% 
  filter(ID %in% RMO_meta$ID) %>% 
  group_by(ID, SpeciesText) %>%       # shouldn't be necessary as there shouldn't be double entries
  summarise(count = sum(count)) %>%   # 
  pivot_wider(names_from = SpeciesText, values_from = count, values_fill = 0)

RMO_spec <- RMO_spec[, !grepl("/", colnames(RMO_spec))] #excluding species aggregates
RMO_spec <- RMO_spec[, !grepl("Obestämd", colnames(RMO_spec))] #excluding records not identified to Species
RMO_spec <- RMO_spec[, !grepl("Nollobservation", colnames(RMO_spec))] #excluding column to store 0s

RMO_spec <- select(RMO_spec, -"NA") %>% ungroup 
```


# Figures

Spatial Figures:

Fig 1A	strata (Esséen et al 2004) needed (SJV)
1. Bakgrundsinfo och resultat på flerfärgade kartor. 

```{r}



NILS <- st_transform(NILS, st_crs(RMO_spat)) %>% 
  st_make_valid(.)

RMO_spat_p <-
  RMO_spat %>% 
  mutate(Data = case_when(county == "SLU" ~ "NILS", 
                          TRUE ~ "Län")) %>% 
  group_by(OBJECTID) %>% 
  st_centroid()


map_cols <- colorRampPalette(brewer.pal(2, "Blues"))(10)

ggplot()+
  geom_sf(data = NILS, aes(fill = StratumNam), alpha = 0.6, size = 0.05)+
  geom_sf_text(data = NILS %>% st_point_on_surface,
               aes(label = OBJECTID), size = 2.5, fontface = "bold")+
  geom_sf(data = RMO_spat_p, size = 0.01, aes(colour = Data))+
  scale_fill_manual(values = map_cols)+
  scale_colour_manual(values = c("darkred", "black"))+
  theme_map()+
  theme(legend.position = "none")+
  blank() +
  north(NILS)+
  scalebar(NILS, dist = 100, dist_unit = "km",
             transform = FALSE, st.size = 2)

ggsave(here("figures", "survey_sites.pdf"))

```
Fig 1B	Distribution of valuable seminatural grasslands needed (SJV)


3. Jämföra transekter och slingor (från de regionala datana). 

4. Jämföra biotoptypernas artantal, antal rödlistade (plus relativ mängd) och artsammansättning (diagram och multivariata matrix).  

5. Sammanvägda trender med TRIM-index för fjärilsgrupper, artnivå i hela regionen och länsvis. 

6. Med multivariata analyser (typ PCA) belysa A) likheter och skillnader i artsammansättningen på län och objektsnivå samt i de olika biotoptyperna (slingsegmenten). B) Vilka parametrar är viktiga för de olika arterna? (nektarrikedom, veg höjd…). 

7. GI-karta – GIS-analys som kombinerar avstånd mellan ängs- och betesmarksobjekt med artrikedom samt rödlistade arter i provpunkterna (LST Örebro – Helena Rygne skall kontaktas för att efterlikna deras körningar). 

8. Lämpliga styrkeanalyser (där hänsyn tas till t.ex.: Signifikansnivån, Effektstorlek., Variationen hos responsvariabeln, Provstorlek och Styrkan).  




check Analysfrågar svar final for how to analyse

## Artrikedom per inventerat objekt

2. Tabeller och diagram med artrikedom per inventerat objekt samt individrikedom (Medelvärden med 95 % CI och totalt antal arter) uppdelat på hela regionen, län och objektsvis. (objekten är stickprov). 

total species richness and richness of grassland specialist (map)

separately between slingor and transects --> check methods previous report 

seperatly between periods 

Method text from the 2015 report:

>Artrikedom och täthet av individer beräknades både för transekter i hagmarkerna och för slingsegmenten i de omgivande markerna. Tätheten utrycks som antal individer/100 m. Eftersom transekterna hade olika längd och därmed var svåra att direkt jämföra med avseende på artrikedom (längre transekter ger fler arter) användes transekter och slingsegment med en längd mellan 100-400 m för dessa beräkningar. Medelvärdet för antal arter i en biotoptyp är således medelvärdet av alla segment som var 100-400 m av respektive biotoptyp. 

>För betesmarkerna gäller samma sak men proceduren var något annorlunda. Här valdes slumpmässigt ut transekter från varje område till dess längd motsvarade medelvärdet för slingsegmentens längder. För varje område gjordes denna procedur om 100 gånger, och sedan användes medelvärdet för alla områden.

>Totalt antal arter per biotop uppskattades också statistiskt med metoden Chao2 och programmet EstimateS. 


Total number of species per Site

I follow request to only include transects but use data from Län & SLU

In the previous report, species richness was standardized by keeping only segments with a length >100m & <400m). However, this 

(1) excludes transect without length information
(2) when I sum the transect segments (per site) the lengths are longer than this range
(3) I could take the middle 50% (25th-75th quantile) of the length but it's not a great solution (and still a factor of 3 between different transects)

--> It seems like we want site totals instead, but that doesn't pair well with the length filtering of segments above (we could keep one segment per site or all, depending on their lengths); also sites have been visited more or less often in the different rounds <--

Instead I will try to pool over sites *and* dates (within each round). 
(1) We can use all data (regardless if length is known)
(2) we standardise by sample completeness (based on total number of observed ind. and the frequency distribution of species)


pool transect and dates
```{r}

RMO_meta_pooled <- 
  RMO_meta %>% 
  filter(SiteType == "T") %>% #keep only trasnect
  select(OBJECTID, round, ID, County_text, county) %>% 
  group_by(County_text, county, OBJECTID, round) %>% 
  summarise(ID = list(unique(ID))) %>% # get a list of what unique transect x date x round ID should be grouped
  mutate(p_ID = paste(OBJECTID, round, sep = "_"))

RMO_spec_pooled <- 
RMO_meta_pooled %>% 
  pull(ID) %>%
  lapply(., function(x) {
    RMO_spec %>% 
    filter( ID %in% x) %>%
      select(-ID) %>% 
      summarise_all(sum)
      }) %>% 
  bind_rows()
```



I use coverage based interpolation - extrapolation based on abundance data to estimate total species richness at each site in each round - and to compare sites at equal coverage

I use the iNEXT package for this

```{r}
# transform into matrix
RMO_spec_pooled <- 
  as.matrix(RMO_spec_pooled)

rownames(RMO_spec_pooled) <- RMO_meta_pooled$p_ID

#calculate estimated richness (and CI) and different levels of sample size (rarefaction) and completeness
out <- iNEXT::iNEXT(t(RMO_spec_pooled), 
                    datatype = "abundance", 
                    q = c(0),
                    nboot = 1 #increase for better estimate
                    )
```


we can look at the distribution of sample completeness across sites and rounds

We see that most sites have actually quite decent sample completeness, of >60 and even > 70 or > 80%

We exclude site with a sample completeness < 60% as the extrapolated diversity from low completeness sites is uncertain and only few sites have this low 
```{r}
out$DataInfo %>% 
  rename("p_ID" = site) %>% 
  left_join(RMO_meta_pooled) %>% 
  select(-ID) %>% 
  ggplot(aes(x = SC))+
  geom_histogram()+
  facet_wrap(~round)
```

we can look at abundance based and sample completeness based rarefaction curves to visualize how complete the assessments are
```{r}
RMO_out <- 
out$iNextEst %>% bind_rows(.id = "p_ID") %>% 
  left_join(select(RMO_meta_pooled, -ID)) %>% 
  mutate(county = case_when(county != "SLU" ~ "LänStyr.",
                            TRUE ~ county))

ggplot(RMO_out, aes(x = m, y = qD, group = OBJECTID, colour=county))+
  geom_line(size = 0.2)+
  geom_point(data = . %>% filter(method == "observed"), size = 0.1, colour = "black")+
  facet_grid(round~County_text)+
  scale_x_continuous(limits = c(0,400))

ggplot(RMO_out, aes(x = SC, y = qD, group = OBJECTID, colour=county))+
  geom_line(size = 0.2)+
  geom_point(data = . %>% filter(method == "observed"), size = 0.1, colour = "black")+
  facet_grid(round~County_text)
```


plot estimated richness 
```{r}
p_ID_excl <- 
RMO_out %>% 
  filter(method == "observed") %>% 
  filter(SC < 0.6) %>% 
  pull(p_ID)


map_cols <- colorRampPalette(brewer.pal(2, "Blues"))(10)

RMO_est <- 
  RMO_spec_pooled %>%
  split(., row(.)) %>% 
  `names<-`(rownames(RMO_spec_pooled)) %>% 
  lapply(., function(x){iNEXT::estimateD(x, base = "coverage", level = 0.95, conf = NULL)}) %>% 
  suppressWarnings() %>%
  bind_rows(.id = "p_ID") %>% 
  left_join(RMO_meta_pooled) 

p1 <- 
RMO_spat %>% 
  group_by(OBJECTID) %>% 
  st_centroid() %>% 
  left_join(RMO_est, .) %>% 
  st_as_sf() %>% 
  ggplot()+
  geom_sf(data = filter(NILS, OBJECTID < 7), 
          aes(fill = StratumNam), alpha = 0.6, 
          size = 0.05,  show.legend = F)+
  geom_sf(size = 1, aes(colour = `q = 0`), alpha = 0.6)+
  facet_wrap(~round)+
  scale_colour_viridis_c(option = "turbo")+
  scale_fill_manual(values = map_cols)+
   theme_map()+
  # theme(legend.position = "right")+
  # blank() +
  # north(filter(NILS, OBJECTID < 7))+
  # scalebar(filter(NILS, OBJECTID < 7),
  #          dist = 100, dist_unit = "km",
  #            transform = FALSE, st.size = 2)+
  ggtitle("average S, all species")+
  theme(legend.position = "bottom")
  
ggsave(here("figures", "estimated_richness_all_spec.pdf"), p1)

p1


```


Same as above for Species defines as grassland specialist

Is it valid to do that? Can I subset data to specialist and then extrapolate N of expected specialist?

**change colour scale**

```{r}

grassland <- 
  Spec_cat %>% 
  filter(Grassland_specialist) %>% 
  pull(SpeciesText)

RMO_spec_pooled_grass <- RMO_spec_pooled[,colnames(RMO_spec_pooled) %in% grassland]
RMO_spec_pooled_grass <- RMO_spec_pooled_grass[rowSums(RMO_spec_pooled_grass) > 0,]

RMO_est_grass <- 
  RMO_spec_pooled_grass %>%
  split(., row(.)) %>% 
  `names<-`(rownames(RMO_spec_pooled_grass)) %>% 
  lapply(., function(x){iNEXT::estimateD(x, base = "coverage", level = 0.95, conf = NULL)}) %>% 
  suppressWarnings() %>%
  bind_rows(.id = "p_ID") %>% 
  left_join(RMO_meta_pooled) 

p2 <- 
RMO_spat %>% 
  group_by(OBJECTID) %>% 
  st_centroid() %>% 
  left_join(RMO_est_grass, .) %>% 
  st_as_sf() %>% 
  ggplot()+
  geom_sf(data = filter(NILS, OBJECTID < 7), 
          aes(fill = StratumNam), alpha = 0.6, 
          size = 0.05,  show.legend = F)+
  geom_sf(size = 1, aes(colour = `q = 0`), alpha = 0.6)+
  facet_wrap(~round)+
  scale_colour_viridis_c(option = "turbo")+
  scale_fill_manual(values = map_cols)+
   theme_map()+
  # theme(legend.position = "right")+
  # blank() +
  # north(filter(NILS, OBJECTID < 7))+
  # scalebar(filter(NILS, OBJECTID < 7),
  #          dist = 100, dist_unit = "km",
  #            transform = FALSE, st.size = 2)+
  ggtitle("estimated number of grassland specialist")+
  theme(legend.position = "bottom")
  
ggsave(here("figures", "estimated_richness_grassland.pdf"), p2)

p2
```


## Antal individer /art/år

vertical barplot of total number of individuals by species (for all species), by round and type

(should it be average per year?)

**change order to have segment next to each other faceting**
```{r}

RMO_tot_ind <- 
  RMO_meta %>% 
  select(SiteType, round, ID) %>% 
  group_by(round, SiteType) %>% 
  summarise(ID = list(unique(ID))) %>% # get a list of what unique transect x date x round ID should be grouped
  mutate(p_ID = paste(round, SiteType, sep = "_"))

RMO_spec_ind <- 
RMO_tot_ind %>% 
  pull(ID) %>%
  lapply(., function(x) {
    RMO_spec %>% 
    filter( ID %in% x) %>%
      select(-ID) %>% 
      summarise_all(sum)
      }) %>% 
  bind_rows()

spec_other <- names(colSums(RMO_spec_ind)[colSums(RMO_spec_ind) < 50])

spec_order <- names(sort(colSums(RMO_spec_ind)))
spec_order <- spec_order[!spec_order %in% spec_other]
spec_order <- c("Other", spec_order)


p3 <- 
RMO_spec_ind %>% 
  mutate(p_ID = RMO_tot_ind$p_ID) %>% 
  pivot_longer(!one_of("p_ID"), names_to = "Species", values_to = "Abundance") %>% 
  left_join(select(RMO_tot_ind, -ID)) %>% 
  mutate(SiteType = ifelse(SiteType == "T", "Transects", "Segments"),
         round = ifelse(round == "round_1", "Round 1", "Round 2")) %>% 
  mutate(Species = ifelse(Species %in% spec_other, "Other", Species)) %>% 
  mutate(Species = factor(Species, levels = spec_order)) %>% 
  ggplot(aes(x = Species, y = Abundance))+
  geom_bar(stat = "identity")+
  facet_wrap(~SiteType*round, nrow = 1)+
  coord_flip()
  #theme_excel_new()

p3

ggsave(here("figures", "N°of_Ind_per_year.pdf"), p3, height = 8, width = 8)
```

Q: which year? (Are there several?)

## Enkel jämförelse – antal arter i transekter och slingor 

(obs bara regionala data). Uppdelat transekter och slingor för att kunna jmf de två perioderna?

Not sure what this is supposed to tell us. The transects and Slingor have different breadth and different lengths...

Should I filter for those with length and standardize per area? (still not super useful as Species~are realtionship is not linear)

```{r}

RMO_p5 <- 
RMO_meta %>% 
  filter(county != "SLU") %>% #exclude SLU
  select(round, OBJECTID, SiteType, date, ID) %>%
  distinct() %>% 
  left_join(RMO_spec) %>% 
  group_by(round, OBJECTID, SiteType, date) %>% 
  summarise(across(one_of(colnames(RMO_spec[,-1])), sum), .groups = "drop") %>% #sum over segemnt on same date
  mutate(across(one_of(colnames(RMO_spec[,-1])), function(x) x>0)) %>%  #code as p/a
  mutate(Richness = rowSums(.[,colnames(RMO_spec[,-1])])) %>% # richness per visit
  select(!one_of(colnames(RMO_spec[,-1]))) %>% 
  group_by(round, OBJECTID, SiteType) %>% 
  summarise(mean_rich = mean(Richness))
  
  
RMO_p5 %>% 
  mutate(SiteType = ifelse(SiteType == "T", "Transects", "Segments"),
         round = ifelse(round == "round_1", "Round 1", "Round 2")) %>% 
  ggplot(aes(x = mean_rich))+
  geom_histogram(binwidth = 3, aes(fill = SiteType), colour = "black", show.legend = FALSE)+
  facet_grid(round~SiteType)+
  theme_excel_new()+
  scale_fill_brewer(palette = "Set1")

ggsave(here("figures", "N_spec_per_observation.pdf"))


```


Histogram of number of species / location, divided between "slingor" and "transektor"

## Biotoptypernas medel artantal

Mean number of species by biotope type (only for Län and only for Slingor)

```{r}
RMO_p6 <-  
RMO_meta %>% 
  filter(county != "SLU") %>% #exclude SLU
  filter(SiteType != "T") %>% 
  filter(!is.na(biotope_text)) %>% 
  select(ID, round, biotope_text) %>% 
  distinct() %>% 
  left_join(RMO_spec) %>% 
  group_by(round, biotope_text) %>% 
  summarise(across(one_of(colnames(RMO_spec[,-1])), sum), .groups = "drop")

RMO_mat_bio <- 
RMO_p6 %>% 
  select(-round, -biotope_text) %>% 
  as.matrix() %>% 
  `rownames<-`(paste(RMO_p6$round, RMO_p6$biotope_text, sep = "__")) 

RMO_out_bio <- 
  RMO_mat_bio %>%
  t() %>% 
  iNEXT(datatype = "abundance", q = c(0),
                    nboot = 1) 

RMO_out_bio$iNextEst %>% 
  bind_rows(.id = "ID") %>% 
  separate(ID, into = c("round", "biotope_text"), sep = "__") %>% 
  ggplot(aes(x = m, y = qD, linetype = as.factor(order), colour = round))+
  geom_line()+
  geom_point(data = . %>% filter(method == "observed"), size = 0.1, colour = "black")+
  facet_wrap(~biotope_text)+
  scale_x_continuous(limits = c(0,5000))

ggsave(here("figures", "richness_by_biotope_type_rarecurves.pdf"))
```

**just show richnes**

```{r}

RMO_bio_div <- 
RMO_mat_bio %>% 
  split(., row(.)) %>% 
  `names<-`(rownames(RMO_mat_bio)) %>% 
  lapply(., function(x){iNEXT::estimateD(x, base = "coverage", level = 0.99, conf = 0.95)}) %>% 
  bind_rows(.id = "ID")

RMO_bio_div <- 
  RMO_bio_div %>% 
  separate(ID, into = c("round", "biotope_text"), sep = "__")


bio_order <- 
  RMO_bio_div %>% 
  filter(order == 0) %>% 
  group_by(biotope_text) %>% 
  summarise(qD = mean(qD)) %>% 
  arrange(desc(qD)) %>% 
  pull(biotope_text)

bio_obs <- 
  RMO_out_bio$iNextEst %>% 
  bind_rows(.id = "ID") %>% 
  separate(ID, into = c("round", "biotope_text"), sep = "__") %>% 
  filter(method == "observed") %>% 
  filter(biotope_text != "Övriga")
    
RMO_bio_div %>%
  filter(order == 0) %>% 
  filter(biotope_text != "Övriga") %>% 
  mutate(biotope_text = factor(biotope_text, levels = bio_order)) %>% 
  ggplot(aes(x = biotope_text, y = qD, ymin = qD.LCL, ymax = qD.UCL, colour = round))+
  geom_pointrange(fatten = 1, position = position_dodge(width = 0.8))+
  geom_point(data = bio_obs, size = 1,alpha = 0.5,
             position = position_dodge(width = 0.8))+
  #facet_grid(~order, scales = "free_y")+
  scale_y_continuous(limits = c(0, NA))+
  #coord_flip()+
  scale_colour_brewer(palette = "Set1")+
  labs(y = "Richness", x = "")+
  theme_bw()+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))
  
ggsave(here("figures", "richness_by_biotope_type.pdf"))

```

## Biotoptypernas artantal och relativa mängd

Total number of species observed by biotoptyp, coloured sep if redlisted, compared to total area of that biotoptyp

--> double axis chart, needs different visualization

```{r}

biotop_frac <- 
RMO_meta %>% 
  filter(county != "SLU") %>% #exclude SLU
  filter(SiteType == "S") %>% 
  filter(!is.na(biotope_text)) %>% 
  filter(!is.na(length)) %>% 
  select(OBJECTID, SiteID, segment, length, biotope_text, round) %>% 
  distinct() %>% 
  group_by(OBJECTID, SiteID, segment, round) %>% 
  slice(1) %>% 
  group_by(biotope_text, round) %>% 
  summarise(length = sum(length)) %>% 
  group_by(round) %>% 
  mutate(length = length/sum(length)*100)

redlist_richness <- 
 RMO_meta %>% 
  filter(county != "SLU") %>% #exclude SLU
  filter(SiteType == "S") %>% 
  filter(!is.na(biotope_text)) %>% 
  select(round, OBJECTID, SiteType, date, ID, biotope_text) %>%
  distinct() %>% 
  left_join(RMO_spec) %>% 
  group_by(round, biotope_text) %>% 
  summarise(across(one_of(colnames(RMO_spec[,-1])), sum), .groups = "drop") %>% 
  mutate(across(one_of(colnames(RMO_spec[,-1])), function(x) x>0)) %>%  #code as p/a
  pivot_longer(one_of(colnames(RMO_spec[,-1])), names_to = "Species", values_to = "presence") %>% 
  mutate(Redlisted = case_when(Species %in% filter(Spec_cat, Redlisted)$SpeciesText ~ TRUE,
         TRUE ~ FALSE)) %>% 
   group_by(round, biotope_text, Redlisted) %>% 
   summarise(Richness = sum(presence)) %>% 
  pivot_wider(names_from = Redlisted, values_from = Richness) %>% 
  rename("Redlisted" = "TRUE", "Richness" = "FALSE")%>% 
  mutate(Richness = Redlisted + Richness) 
 


redlist_richness %>% 
  left_join(biotop_frac) %>% 
  pivot_longer(cols = one_of("Richness", "Redlisted", "length"), 
               names_to = "Var", values_to = "val") %>% 
  mutate(biotope_text = factor(biotope_text, levels = bio_order)) %>% 
  mutate(Var = case_when(Var == "length" ~ "% of each habitat",
                         Var == "Richness" ~ "obs. tot. Richness",
                         Var == "Redlisted" ~ "obs. Redlisted")) %>% 
  filter(biotope_text != "Övriga") %>% 
  ggplot(aes(x = biotope_text, y = val, fill = round))+
  facet_grid(~Var, scales = "free")+
  geom_bar(stat = "identity", position = "dodge")+
  theme_bw()+
  scale_fill_brewer(palette = "Set1")+
  coord_flip()+
  labs(x = "", y = "")+
  theme(legend.position = "bottom")

ggsave(here("figures", "observed_richness_by_biotope.pdf"))
```

```{r}

redlist_richness %>% 
  left_join(biotop_frac) %>% 
  filter(biotope_text != "Övriga") %>% 
  ggplot(aes(x = length, y = Richness, colour = round))+
  geom_point()+
  theme_bw()+
  scale_colour_brewer(palette = "Set1")+
  labs(y = "observed Richness", x = "fraction of habitat")+
  theme(legend.position = "bottom")

```
##  Medeltätheter per habitat, Bastardsvärmare, Rödlistade

Butterfly density per habitat per 100 m

```{r}
RMO_p8 <- 
RMO_meta %>% 
  filter(county != "SLU") %>% #exclude SLU
  filter(SiteType == "S") %>% 
  filter(!is.na(biotope_text)) %>% 
  filter(!is.na(length)) %>% 
  select(ID, OBJECTID, SiteID, segment, length, biotope_text, round) %>% 
  distinct() %>% 
  group_by(ID, OBJECTID, SiteID, segment, round) %>% 
  slice(1) %>% 
  left_join(RMO_spec) %>% 
  group_by(OBJECTID, round, biotope_text) %>% 
  summarise(across(one_of(c("length", colnames(RMO_spec[,-1]))), sum), .groups = "drop") %>% 
  pivot_longer(one_of(colnames(RMO_spec[,-1])), names_to = "Species", values_to = "Abundance") %>% 
  group_by(OBJECTID, round, biotope_text) %>% 
  summarise(Abundance = sum(Abundance),
            length = unique(length)) %>% 
  mutate(density = Abundance / (length/100)) 

RMO_p8 %>% 
  mutate(biotope_text = factor(biotope_text, levels = bio_order)) %>% 
  filter(biotope_text != "Övriga") %>%
    ggplot(aes(x = biotope_text, y = density, colour = round))+
    geom_point(position = position_jitterdodge(jitter.width = 0.2, jitter.height =  0,
                                               dodge.width = 0.8), size = 0.5, alpha = 0.5)+
    stat_summary(fun.data = "mean_cl_boot", fatten = 0.8, size = 0.3, colour = "black",
                 position = position_dodge(width = 0.8), aes(group = round))+
    theme_bw()+
    scale_colour_brewer(palette = "Set1")+
    coord_flip()+
    labs(x = "", y = "")+
    theme(legend.position = "bottom")+
    scale_y_continuous(limits = c(0,15))
  
ggsave(here("figures", "density_per_100m_per_biotop.pdf"))
```

only redlisted
```{r}

Redlisted <- 
  Spec_cat %>% 
  filter(Redlisted) %>% 
  filter(SpeciesText %in% colnames(RMO_spec)) %>% 
  pull(SpeciesText)
  


RMO_p8.1 <- 
RMO_meta %>% 
  filter(county != "SLU") %>% #exclude SLU
  filter(SiteType == "S") %>% 
  filter(!is.na(biotope_text)) %>% 
  filter(!is.na(length)) %>% 
  select(ID, OBJECTID, SiteID, segment, length, biotope_text, round) %>% 
  distinct() %>% 
  group_by(ID, OBJECTID, SiteID, segment, round) %>% 
  slice(1) %>% 
  left_join(select(RMO_spec, one_of(c("ID", Redlisted)))) %>% 
  group_by(OBJECTID, round, biotope_text) %>% 
  summarise(across(one_of(c("length", Redlisted)), sum), .groups = "drop") %>% 
  pivot_longer(one_of(Redlisted), names_to = "Species", values_to = "Abundance") %>% 
  group_by(OBJECTID, round, biotope_text) %>% 
  summarise(Abundance = sum(Abundance),
            length = unique(length)) %>% 
  mutate(density = Abundance / (length/100)) 

RMO_p8.1 %>% 
  mutate(biotope_text = factor(biotope_text, levels = bio_order)) %>% 
  filter(biotope_text != "Övriga") %>%
    ggplot(aes(x = biotope_text, y = density, colour = round))+
    geom_point(position = position_jitterdodge(jitter.width = 0.2, jitter.height =  0,
                                               dodge.width = 0.8), size = 0.5, alpha = 0.5)+
    stat_summary(fun.data = "mean_cl_boot", fatten = 0.8, size = 0.3, colour = "black",
                 position = position_dodge(width = 0.8), aes(group = round))+
    theme_bw()+
    scale_colour_brewer(palette = "Set1")+
    coord_flip()+
    labs(x = "", y = "")+
    theme(legend.position = "bottom")+
    scale_y_log10()

ggsave(here("figures", "density_redlisted_per_100m_per_biotop.pdf"))
```



## Samband mängd nektarkällor och för artantal och individantal

**split by round**

species abundance vs nectar abundance

```{r}
RMO_p9.1 <- 
RMO_meta %>% 
  filter(SiteType == "T") %>% 
  filter(!is.na(length)) %>% 
  filter(!is.na(per_mill_flowers)) %>% 
  select(ID, date, OBJECTID, SiteID, segment, length,per_mill_flowers, round) %>% 
  distinct() %>% 
  group_by(ID) %>% 
  slice(1) %>%  #if there are double entries for the same segment on the same date we chosse the first
  left_join(RMO_spec) %>% 
  pivot_longer(one_of(colnames(RMO_spec[,-1])), names_to = "Species", values_to = "Abundance") %>% 
  group_by(ID) %>% 
  summarise(Abundance = sum(Abundance),
            length = unique(length),
            per_mill_flowers = unique(per_mill_flowers)) %>% 
  mutate(Density = Abundance/(length/100))

ggplot(RMO_p9.1, aes(x = per_mill_flowers, y = Density))+
  geom_point(size = 0.8, alpha = 0.4, colour = "deepskyblue4")+
  geom_smooth(method = "lm", se = F, size = 0.8, colour = "darkred")+
  scale_y_log10()+
  scale_x_log10()+
  theme_bw()+
  labs(x = "‰ flowers", y = "Abundance / 100m")

ggsave(here("figures", "flower_dens_vs_butterfly_dens.pdf"))
```

barplot
```{r}
RMO_p9.1 %>% 
  mutate(per_mill_flowers_cat = 
           case_when(per_mill_flowers == 0 ~ "0",
                     per_mill_flowers <= 1 ~ "0-1",
                     per_mill_flowers > 1 & per_mill_flowers <= 2 ~ "1-2",
                     per_mill_flowers > 2 ~ ">2")) %>% 
  mutate(per_mill_flowers_cat = factor(per_mill_flowers_cat, levels = c("0", "0-1", "1-2", ">2"))) %>% 
  group_by(per_mill_flowers_cat) %>% 
  summarise(mean_cl_boot(Density)) %>% 
  ggplot(aes(x = per_mill_flowers_cat, y = y, ymin = ymin, ymax = ymax))+
  geom_pointrange(fatten = 0.5)+
  scale_y_continuous(limits = c(0,NA))+
  theme_bw()+
  labs(x = "‰ flowers", y = "Abundance / 100m")

ggsave(here("figures", "flower_dens_vs_butterfly_dens[grouped].pdf"))
```


richness vs nectar abundance
```{r}

RMO_p9.2 <- 
RMO_meta %>% 
  filter(SiteType == "T") %>% 
  filter(!is.na(length)) %>% 
  filter(!is.na(per_mill_flowers)) %>% 
  select(ID, date, OBJECTID, SiteID, segment, length,per_mill_flowers, round) %>% 
  distinct() %>% 
  group_by(ID) %>% 
  slice(1) %>%  #if there are double entries for the same segment on the same date we chosse the first
  filter(length > 50 && length < 200) %>% 
  left_join(RMO_spec) %>% 
  pivot_longer(one_of(colnames(RMO_spec[,-1])), names_to = "Species", values_to = "Abundance") %>% 
  mutate(Abundance = ifelse(Abundance > 0, 1, 0)) %>% 
  group_by(ID) %>% 
  summarise(Richness = sum(Abundance),
            length = unique(length),
            per_mill_flowers = unique(per_mill_flowers)) 

ggplot(RMO_p9.2, aes(x = per_mill_flowers, y = Richness))+
  geom_point(size = 0.8, alpha = 0.4, colour = "deepskyblue4")+
  geom_smooth(method = "lm", se = F, size = 0.8, colour = "darkred")+
  scale_y_log10()+
  scale_x_log10()+
  theme_bw()+
  labs(x = "‰ flowers")

ggsave(here("figures", "flower_dens_vs_richness.pdf"))

```


```{r}
RMO_p9.2 %>% 
  mutate(per_mill_flowers_cat = 
           case_when(per_mill_flowers == 0 ~ "0",
                     per_mill_flowers <= 1 ~ "0-1",
                     per_mill_flowers > 1 & per_mill_flowers <= 2 ~ "1-2",
                     per_mill_flowers > 2 ~ ">2")) %>% 
  mutate(per_mill_flowers_cat = factor(per_mill_flowers_cat, levels = c("0", "0-1", "1-2", ">2"))) %>% 
  group_by(per_mill_flowers_cat) %>% 
  summarise(mean_cl_boot(Richness)) %>% 
  ggplot(aes(x = per_mill_flowers_cat, y = y, ymin = ymin, ymax = ymax))+
  geom_pointrange(fatten = 0.5)+
  scale_y_continuous(limits = c(0,NA))+
  theme_bw()+
  labs(x = "‰ flowers", y = "Richness")

ggsave(here("figures", "flower_dens_vs_richness[grouped].pdf"))
```


## Samband vegetationshöjd och  artantal och individantal

not sure how to do this

Veg height is given in how much of each of three categories. Would need to make a regression analysis for that. 

```{r}

RMO_p10.1 <- 
RMO_meta %>% 
  filter(SiteType == "T") %>% 
  filter(!is.na(length)) %>%
  select(ID, starts_with("perc_veg_")) %>% 
  distinct() %>% 
  group_by(ID) %>% 
  slice(1) %>%  #if there are double entries for the same segment on the same date we chosse the first
  pivot_longer(cols = starts_with("perc_veg_"), names_to = "veg_cat", values_to = "veg_height") %>% 
  group_by(ID) %>% 
  filter(!all(is.na(veg_height)))
  select(ID, date, OBJECTID, SiteID, segment, length,per_mill_flowers, round) %>% 
  distinct() %>% 
 
  left_join(RMO_spec) %>% 
  pivot_longer(one_of(colnames(RMO_spec[,-1])), names_to = "Species", values_to = "Abundance") %>% 
  group_by(ID) %>% 
  summarise(Abundance = sum(Abundance),
            length = unique(length),
            per_mill_flowers = unique(per_mill_flowers)) %>% 
  mutate(Density = Abundance/(length/100))

ggplot(RMO_p9.1, aes(x = per_mill_flowers, y = Density))+
  geom_point(size = 0.8, alpha = 0.4, colour = "deepskyblue4")+
  geom_smooth(method = "lm", se = F, size = 0.8, colour = "darkred")+
  scale_y_log10()+
  scale_x_log10()+
  theme_bw()+
  labs(x = "‰ flowers", y = "Abundance / 100m")

```


## Typiska arter för resp biotop (slingssegmenten)

(should we do some multivarte anlaysis here to see what species are sign more abundant in certain biotop type?)

```{r}
RMO_p11 <- 
RMO_meta %>% 
  filter(county != "SLU") %>% #exclude SLU
  filter(SiteType == "S") %>% 
  filter(!is.na(biotope_text)) %>% 
  filter(!is.na(length)) %>% 
  select(ID, OBJECTID, SiteID, segment, length, biotope_text) %>% 
  distinct() %>% 
  group_by(ID, OBJECTID, SiteID, segment) %>% 
  dplyr::slice(1) %>% 
  left_join(RMO_spec) %>% 
  group_by(biotope_text) %>% 
  summarise(across(one_of( colnames(RMO_spec[,-1])), sum), .groups = "drop") %>% 
  pivot_longer(one_of(colnames(RMO_spec[,-1])), names_to = "Species", values_to = "Abundance") %>% 
  group_by(Species) %>% 
  filter(sum(Abundance) > 0) %>% 
  group_by(biotope_text) %>% 
  mutate(Abundance = Abundance / sum(Abundance, na.rm = TRUE)) %>% #standardise by total observations in bitopoe
  group_by(Species) %>% 
  mutate(rel_abund = Abundance / sum(Abundance, na.rm = TRUE)) #get rel abundance 

p11 <- 
RMO_p11 %>% 
  mutate(biotope_text = factor(biotope_text, levels = bio_order)) %>% 
  filter(biotope_text != "Övriga") %>%
    ggplot(aes(x = biotope_text, y = Species, fill = rel_abund))+
    geom_tile()+
    theme_bw()+
    scale_fill_viridis_c()+
  theme(axis.text.x = element_text(angle  = -90, hjust = 0))

p11

ggsave(here("figures", "typical_species.pdf"), width = 8, height = 10)
```



## Trend analysis 

Is this dealt with?

## Jmf av artsammansättning mellan länen och de ingående objekten

```{r}
RMO_12 <- 
RMO_meta %>% 
  #filter(SiteType == "T") %>% 
  select(ID, OBJECTID, County_text) %>% 
  distinct() %>% 
  group_by(ID) %>% 
  dplyr::slice(1) %>% 
  left_join(RMO_spec) %>% 
  group_by(OBJECTID, County_text) %>% 
  summarise(across(one_of( colnames(RMO_spec[,-1])), sum), .groups = "drop") 

RMO_12_meta <- 
  RMO_12 %>% 
  select(County_text, OBJECTID) %>% 
  mutate(ID = paste(County_text, OBJECTID)) %>% 
   as.data.frame() %>% 
  `rownames<-`(.$ID)

RMO_12_mat <- 
  RMO_12 %>% 
  select(-County_text, -OBJECTID) %>% 
  as.matrix() %>% 
  `rownames<-`(RMO_12_meta$ID)

RMO_12_ps = phyloseq(otu_table(RMO_12_mat, taxa_are_rows = FALSE),
                     sample_data(RMO_12_meta))

RMO_12_dd <- phyloseq_to_deseq2(RMO_12_ps, ~ County_text)

RMO_12_dd = DESeq(RMO_12_dd, test="Wald", fitType="parametric", sfType = "poscounts")

RMO_12_vst <- getVarianceStabilizedData(RMO_12_dd) %>% t()

RMO_12_vst[RMO_12_vst < 0] <- 0

RMO_12_vst_d <- vegdist(RMO_12_vst, "bray")
  
RMO_12_nmds <- metaMDS(RMO_12_vst, autotransform = FALSE, sep = 50)

RMO_12_points <- 
RMO_12_nmds$points %>% 
  as_tibble(rownames = "ID") %>% 
  left_join(RMO_12_meta)

RMO_hull_12 <- RMO_12_points %>% 
  group_by(County_text) %>% 
  dplyr::slice(chull(MDS1, MDS2))
  
ggplot(RMO_12_points, aes(x = MDS1, y = MDS2, colour = County_text))+
  geom_point()+
  geom_polygon(data = RMO_hull_12, fill = NA, size = 0.2)+
  theme_bw()+
  scale_colour_brewer(palette = "Set1")
  
#stressplot(RMO_12_nmds)
```



## Jmf artsammansättningen för de olika biotoptyperna (sling segmenten)

its messy as there are a lot of habitats...

Also, I cannot plot the single sites as even after pooling multiple visits they have just too few observations per habitat to calculate a realistic beta-diversity. 

Therefore I pool habitats by county, but that (of course) removes a lot of variability. 

```{r}
RMO_13 <- 
RMO_meta %>% 
  select(ID, OBJECTID, biotope_text, SiteType, County_text) %>% 
  filter(SiteType == "S") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  dplyr::slice(1) %>%
  filter(!is.na(biotope_text)) %>% 
  left_join(RMO_spec) %>% 
  group_by(County_text, biotope_text) %>% 
  summarise(across(one_of( colnames(RMO_spec[,-1])), sum), .groups = "drop") 

RMO_13_meta <- 
  RMO_13 %>% 
  select(County_text, biotope_text) %>% 
  mutate(ID = paste(County_text, biotope_text)) %>% 
  as.data.frame() %>% 
  `rownames<-`(.$ID)

RMO_13_mat <- 
  RMO_13 %>% 
  select(-County_text, -biotope_text) %>% 
  as.matrix() %>% 
  `rownames<-`(RMO_13_meta$ID)

RMO_13_mat <- RMO_13_mat[rowSums(RMO_13_mat) > 0,]

RMO_13_ps = phyloseq(otu_table(RMO_13_mat, taxa_are_rows = FALSE),
                     sample_data(RMO_13_meta))

RMO_13_dd <- phyloseq_to_deseq2(RMO_13_ps, ~ biotope_text)

RMO_13_dd = DESeq(RMO_13_dd, test="Wald", fitType="parametric", sfType = "poscounts")

RMO_13_vst <- getVarianceStabilizedData(RMO_13_dd) %>% t()

RMO_13_vst[RMO_13_vst < 0] <- 0

RMO_13_vst <- RMO_13_vst[rowSums(RMO_13_vst) > 0,]

RMO_13_vst_d <- vegdist(RMO_13_vst)
#RMO_13_vst_d_s <- stepacross(RMO_13_vst_d)

RMO_13_nmds <- metaMDS(RMO_13_vst_d, autotransform = FALSE, try = 50)

RMO_13_points <- 
RMO_13_nmds$points %>% 
  as_tibble(rownames = "ID") %>% 
  left_join(RMO_13_meta)

RMO_hull_13 <- RMO_13_points %>% 
  group_by(biotope_text) %>% 
  dplyr::slice(chull(MDS1, MDS2))
  
ggplot(RMO_13_points, aes(x = MDS1, y = MDS2, colour = biotope_text))+
  geom_point()+
  geom_polygon(data = RMO_hull_13, fill = NA)+
  theme_bw()
  scale_colour_brewer(palette = "Set1")
  
  RMO_hull_13_c <- RMO_13_points %>% 
  group_by(County_text) %>% 
  dplyr::slice(chull(MDS1, MDS2))
  
ggplot(RMO_13_points, aes(x = MDS1, y = MDS2, colour = County_text))+
  geom_point()+
  geom_polygon(data = RMO_hull_13_c, fill = NA)+
  theme_bw()+
  scale_colour_brewer(palette = "Set1")
  
```

## Vilka parametrar (nektarrikedom, veghöjd…) viktiga för olika arterna?

what env var are important for different species. Vegan correspondance analysis. 

Q: What env variables do we want to include here?

## Potentialen att hysa gräsmarksfjärilsspecialister

--> check R code Spatial for this and some more spatial analysis




```{r}
RMO_meta %>% 
  mutate(county = ifelse(county != "SLU", County_text, "SLU")) %>% 
  select(county, OBJECTID, SiteType, segment, length) %>% 
  mutate(length = !is.na(length)) %>% 
  group_by(OBJECTID) %>% 
  mutate(N_segments = length(unique(segment))) %>% 
  select(-segment) %>% 
  distinct() %>% 
  group_by(county, SiteType, length) %>% 
  summarise(N_sites = length(unique(OBJECTID)),
            N_segments = round(mean(N_segments)))%>% 
  write_clip


RMO_spat %>% st_crs()
  dplyr::filter(county != "SLU") %>% 
  st_write(., "LST_midpoint.csv", layer_options = "GEOMETRY=AS_XY")
```





