---
title: "ÄoB analysis of Butterflydata"
author: "Fabian Roger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes

---

load packages

```{r}
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
```

load data

```{r}
RMO <- read_excel("RMO_combined_data.xlsx", sheet = "all_observations_clean_no_calc", 
                  col_types = c(
                    "skip", "text", "text", "numeric", "text",
                    "text", "numeric", "text", rep("numeric",5),
                    rep("skip",3), rep("numeric",4),
                    rep("text",7), rep("skip", 16)),
                  n_max = 2000
                  )
```


# Figures

## Artrikedom per inventerat objekt

total species richness and richness of grassland specialist (map)

## Antal individer /art/år

vertical barplot of total number of individuals by species (for all species) 

Q: which year? (Are there several?)

## Enkel jämförelse – antal arter i transekter och slingor 

(obs bara regionala data). Uppdelat transekter och slingor för att kunna jmf de två perioderna?

Histogram of number of species / location, divided between "slingor" and "transektor"

## Biotoptypernas medel artantal + jmf perioderna

Mean N species + CI by biotoptyp

## Biotoptypernas artantal och relativa mängd

Total number of species observed by biotoptyp, coloured sep if redlisted, compared to total area of that biotoptyp

--> double axis chart, needs different visualization







