---
title: "Import and data cleanup"
author: "Fabian Roger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
---

load packages

```{r}
library(dplyr)
library(here)
library(tidyr)
library(readxl)
library(sf)
library(lubridate)
```



## Data description:

+ Länstyrelsen data (county codes other than SLU)

--> biotop only for Länstyrelsen
--> Slingor and Transects, analyse separately unless they have length so we can standardise (see description of wanted analysis)

+ SLU data (collected by SLU)

+ from 2009 to 2018 (doublecheck)
There is different periods 9-13 and 14-18, Analyse separately?

`vadd_tist_klint` only SLU

Spatial data in sheet 'site_geo' EPSG:3021

--> get county info for SLU data from county shapefile and geo-location of routes

Work with Species ID and check for consistent coding
SiteID and Site type --> no gaps, unique ID

## loading the data

We cannot load the data directly as we need to fix them to have a single type per column. 
We therefore import the data as text 
```{r}
#data
RMO <- read_excel("RMO_combined_data_updated.xlsx", sheet = "all_observations_clean_no_calc", 
                  col_types = "text")

#variable description
Var_RMO <- read_excel("RMO_combined_data_updated.xlsx", sheet = 1)

# Var_RMO %>%
#   filter(Variable %in% colnames(RMO)) %>%
#   View(., title = "Variables")

#spatial data
RMO_spat <- read_excel("RMO_combined_data_updated.xlsx", sheet = "site_geo")

# TUVA (AOB) shape-file
TUVA <- read_sf(here("TUVA", "AoB_2019.shp"))
```

## check variables

### `date`

date has been imported as serial number (and transformed to character)

However some dates are still in character date format 

```{r}
# these date have not been imported as numeric (not sure why)
RMO$date[is.na(as.numeric(RMO$date))] %>%
  table()
```

```{r}
# all but 2 follow the %Y-%m-%d format
RMO$date[is.na(as.numeric(RMO$date))] %>%
  unique() %>% 
  as.Date(., format = "%Y-%m-%d")
```

1) 2010-20-08 --> probably %Y-%d-%m, replace by 2010-08-20 before transformation

2) 2018-06-31 --> impossible date, replace by 2018-07-01

3) transform all character 

4) transform excel dates


```{r}
#1
RMO$date[RMO$date == "2010-20-08"] <- "2010-08-20"

#2
RMO$date[RMO$date == "2018-06-31"] <- "2018-07-01"

## OBS !! different origin for mac and windows excel !! ###
## from ?as.Date:
## Excel is said to use 1900-01-01 as day 1 (Windows default) or
## 1904-01-01 as day 0 (Mac default)

## it seems like the origin here is "1899-12-30"

#3 & #4
RMO <- 
RMO %>% 
  mutate(date = case_when(is.na(as.numeric(date)) ~ as.Date(date, format = "%Y-%m-%d"),
                          TRUE ~ as.Date(as.numeric(date), origin = "1899-12-30"))) 

```

### `count`

no problems 

```{r}
as.numeric(RMO$count) %>% is.na %>% sum
```

```{r}
RMO$count <- as.numeric(RMO$count)
```

### `county`

```{r}

RMO$county %>% table()
RMO$CountyCode %>% table()

```

`county` contains letter codes for the counties (Länstyrelsen) and 'SLU' if the data have been recorded by SLU
`CountyCode` contains number code for counties, including for SLU

I add a variable `County_text` with the names of the counties based on the number code

```{r}

County_names <- 
Var_RMO %>% 
  filter(Variable == "CountyCode") %>% 
  pull(Values) %>% 
  strsplit(",") %>% 
  unlist() %>% 
  tibble(var = .) %>% 
  separate(var, into = c("CountyCode", "County_text"), sep = "=") %>% 
  mutate(CountyCode = stringr::str_trim(CountyCode)) 
  

RMO <- RMO %>% 
  left_join(County_names)

```

### `segment``

```{r}
RMO[which(is.na(as.numeric(RMO$segment))),"segment"] %>% table()
```

1) S1-17 --> strip "S"
2) S --> 1
3) A-K --> 1-11
4) drop rest

```{r}
#1
RMO$segment[grepl("S", RMO$segment)] <- gsub("S(\\d+)", "\\1", RMO$segment[grepl("S", RMO$segment)]) 

#2
RMO$segment[RMO$segment == "S"] <- 1

#3
Let_coded <- RMO$segment[grepl(paste0(LETTERS[1:11], collapse = "|"), RMO$segment)]
RMO$segment[grepl(paste0(LETTERS[1:11], collapse = "|"), RMO$segment)] <-  match(Let_coded, LETTERS)

#4
RMO[which(is.na(as.numeric(RMO$segment))),"segment"] <- NA_character_

#transform
RMO$segment <- as.numeric(RMO$segment)
```

### `biotope`

```{r}
RMO$biotope[!is.na(RMO$biotope)][which(is.na(as.numeric(RMO$biotope[!is.na(RMO$biotope)])))] %>% table() 
```

1) remove all entries where habitat is not uniquely identified

2) check if NAs for Slingor (SiteType == "S) can be imputed (Transect do generally not have habitat information, but 14 transect sites seem to have it anyway)


```{r}
#1
RMO$biotope[!is.na(RMO$biotope)][which(is.na(as.numeric(RMO$biotope[!is.na(RMO$biotope)])))] <- NA_character_

#get table for code - name translation
biotope_names <- 
Var_RMO %>% 
  filter(Variable == "biotope") %>% 
  pull(Values) %>% 
  strsplit(",") %>% 
  unlist() %>% 
  tibble(var = .) %>% 
  separate(var, into = c("biotope", "biotope_text"), sep = "=") %>% 
  mutate(biotope = stringr::str_trim(biotope),
         biotope_text = stringr::str_trim(biotope_text)) %>% 
  mutate(biotope_text = case_when(biotope == "16" ~ "Övriga",
                                  TRUE ~ biotope_text))

# add biotope_names to RMO
RMO <- 
left_join(RMO, biotope_names)

#check if biotopes are consistent accross repeated entries
RMO %>% 
  filter(SiteType == "S") %>% 
  filter(!is.na(biotope_text)) %>% 
  select(SiteID, segment, biotope_text, date) %>%
  mutate(date = year(date)) %>% 
  distinct() %>% 
  group_by(SiteID, segment, date) %>% 
  mutate(n_bio = length(unique(biotope_text))) %>% 
  filter(n_bio > 1) %>% 
  #group_by(SiteID, segment, date) %>% 
  #filter(n() == 1 ) %>% 
  arrange(SiteID, segment, date) %>% 
  mutate(biotope_text = paste(biotope_text, " (", date, ")", sep = "")) %>% 
  group_by(SiteID, segment, date) %>% 
  mutate(year = paste("case", 1:n())) %>% 
  select(-n_bio) %>% 
  pivot_wider(names_from = year, values_from = biotope_text) %>% 
  inner_join(distinct(select(RMO, SiteText, SiteID))) %>% 
  relocate("SiteText", .before = "SiteID") %>% View(., "biotopes") 
  #write.table("Segments_biotope_conflicts_same_year.txt", sep = "\t", row.names = FALSE)
```



```{r}
# some biotope types can be imputed 
impute_bio <- 
RMO %>% 
  filter(SiteType == "S") %>% 
  select(SiteID, segment, date, biotope) %>% 
  distinct() %>% 
  group_by(SiteID, segment) %>% 
  filter(TRUE %in% is.na(biotope)) %>% 
  filter(!all(is.na(biotope))) %>% 
  arrange(SiteID, segment, date) 

#not all are uniquely identified, we only impute those that are so
bio_not_same <- 
impute_bio %>% 
  filter(!is.na(biotope)) %>% 
  group_by(SiteID, segment) %>% 
  filter(length(unique(biotope)) > 1)

bio_not_same

bio_same <- 
  impute_bio %>% 
  filter(!is.na(biotope)) %>% 
  group_by(SiteID, segment) %>% 
  filter(length(unique(biotope)) == 1) %>% 
  select(-date) %>% 
  distinct()

#impute
RMO <- 
left_join(RMO, bio_same, by = c("SiteID", "segment")) %>% 
  mutate(biotope = coalesce(biotope.y, biotope.x)) %>%
  select(-biotope.y, -biotope.x)
```

### `length`

```{r}
RMO$length[!is.na(RMO$length)][which(is.na(as.numeric(RMO$length[!is.na(RMO$length)])))]
```

```{r}
RMO %>% 
  select(SiteID, segment, length, date) %>%
  filter(!is.na(length)) %>% 
  group_by(SiteID, segment, length) %>% 
  mutate(n = n()) %>% 
  select(-date) %>% 
  distinct() %>% 
  group_by(SiteID, segment) %>%  
  filter(n() > 1) %>% 
  arrange(SiteID, segment) %>% 
  mutate(length = paste(length, " (", n, ")", sep = "")) %>% 
  select(-n) %>% 
  mutate(n_l = 1:n()) %>% 
  pivot_wider(names_from = n_l, values_from = length) %>% 
  inner_join(distinct(select(RMO, SiteText, SiteID))) %>% 
  relocate("SiteText", .before = "SiteID") #%>% 
  #write.table("Segments_length_conflicts.txt", sep = "\t", row.names = FALSE)
  
```



1) c <- 246
2) i.u. seems to code for missing (inte undersökt?) --> we code it as NA
3) for many segments the length is not the same between dates. This is probably genuine as the length then represents the observed length at that date. However, in a few cases the length is given as 335000 which is an impossible value. We replace it by the recorded length on another day
4) for many segments and transect the length is NA at least at some dates. If the length is always constant, we can impute it. 

```{r}

#1
RMO %>% 
  filter(SiteID == RMO[!is.na(RMO$length) & RMO$length == "c",]$SiteID,
         segment == RMO[!is.na(RMO$length) & RMO$length == "c",]$segment) %>% 
  select(SiteID, segment, date, length)

RMO$length[!is.na(RMO$length) & RMO$length == "c"] <- "246"

#2
RMO$length[!is.na(RMO$length) & RMO$length == "i.u"] <- NA_character_

#3
right_len <- 
RMO %>% 
  filter(SiteID %in% RMO[!is.na(RMO$length) & RMO$length == "335000",]$SiteID) %>% 
  select(SiteID, segment, date, length) %>% 
  distinct() %>% 
  arrange(SiteID, segment, date) %>% 
  group_by(SiteID, segment) %>% 
  summarise(length = unique(length)) %>% 
  filter(length != "335000") 

RMO <- RMO %>% 
left_join(right_len, by = c("SiteID", "segment")) %>%  
  mutate(length = coalesce(length.y, length.x)) %>% 
  select(-length.y, -length.x)

#for one segment there is no date with non-impossible length. We set it to NA
RMO[!is.na(RMO$length) & RMO$length == "335000",]$length <- NA_character_

#4

imputable_length <- 
RMO %>% 
  select(date, SiteID, segment, length) %>% 
  group_by(SiteID, segment) %>% 
  filter(TRUE %in% is.na(length)) %>%
  filter(FALSE %in% is.na(length)) %>% 
  filter(!is.na(length)) %>% 
  select(-date) %>% 
  distinct() %>% 
  group_by(SiteID, segment) %>% 
  filter(n()==1)

RMO <- 
RMO %>% 
left_join(imputable_length, by = c("SiteID", "segment")) %>%  
  mutate(length = coalesce(length.y, length.x)) %>% 
  select(-length.y, -length.x) 
  
#transform to numeric
RMO$length <- as.numeric(RMO$length)

```


### `perc_veg_X` 

```{r}
#perc_veg_0-5 --> needs fixing. Includes comments, symbols ("x") and unrecognized NAs (that are coded as "NA" in raw data)
RMO$`perc_veg_0-5`[!is.na(RMO$`perc_veg_0-5`)][which(is.na(as.numeric(RMO$`perc_veg_0-5`[!is.na(RMO$`perc_veg_0-5`)])))] %>% 
  table() 

#perc_veg_5-15 --> needs fixing. Includes comments, symbols ("x") and unrecognized NAs (that are coded as "NA" in raw data)
RMO$`perc_veg_5-15`[!is.na(RMO$`perc_veg_5-15`)][which(is.na(as.numeric(RMO$`perc_veg_5-15`[!is.na(RMO$`perc_veg_5-15`)])))] %>% table() 

#perc_veg_above_15 --> needs fixing. Includes comments, symbols ("x") and unrecognized NAs (that are coded as "NA" in raw data)
RMO$`perc_veg_above_15`[!is.na(RMO$`perc_veg_above_15`)][which(is.na(as.numeric(RMO$`perc_veg_above_15`[!is.na(RMO$`perc_veg_above_15`)])))] %>% table() 

# should only be present in transects and should ideally sum to 100% between three size classes
# NAs might be 0 if other categories sum to 100
# x's are likely 100 if no x in other categories
```

```{r}
not_100 <- 
RMO %>% 
  filter(SiteType == "T") %>% #only trnasects should have cover data
  select(SiteID,SiteType, segment, date, starts_with("perc_veg")) %>% 
  distinct() %>% 
  rowwise() %>% 
  mutate(perc_veg_100 = sum(as.numeric(c_across(starts_with("perc_veg"))), na.rm = T)) %>% 
  suppressWarnings() %>% 
  filter(perc_veg_100 != 100) %>% 
  distinct() 

# for about 14% of the Site x segment x dates combinations, the prc vegetation does not add up to 100
paste("percentage of Site x segment x dates combination were sum(% veg) is not 100: ", nrow(not_100)/17757*100)

#over 100
not_100 %>% 
  filter(perc_veg_100 > 100) %>% 
  nrow(.) %>% 
  paste("over_100:", .)

#over 110
not_100 %>% 
  filter(perc_veg_100 > 110) %>% 
  nrow(.) %>% 
  paste("over_110:", .)

#all NA
not_100 %>% 
  select(-perc_veg_100) %>% 
  filter(sum(is.na(c_across(starts_with("perc_veg")))) == 3) %>% 
  nrow() %>% 
  paste("allNA:", .)

#coded as x
not_100 %>% 
  filter(`perc_veg_0-5` %in% c("x", "X") | `perc_veg_5-15`  %in% c("x", "X") | `perc_veg_above_15`  %in% c("x", "X")) %>% 
  nrow() %>% 
  paste("coded as X:", .)

#less than 100
not_100 %>% 
  filter(perc_veg_100 > 0 & perc_veg_100 < 100) %>% 
  nrow() %>% 
  paste("less than 100", .)

print("other problems: ")

#other problems
not_100 %>% 
  filter(perc_veg_100 == 0) %>% 
  select(-perc_veg_100) %>% 
  rowwise() %>% 
  filter(sum(is.na(c_across(starts_with("perc_veg")))) < 3) %>%  #not all 3 NA
  filter(all(is.na(match(c("x", "X") ,c(`perc_veg_0-5`, `perc_veg_5-15`, `perc_veg_above_15`))))) %>% 
  pull(`perc_veg_0-5`) %>% table() 
```
Only transects should have data on vegetation cover. Most have, but there are problems:

1) not all cover add to 100. This could be genuine as the observers can count other surfaces.
2) some cover adds to >100 --> check manually
3) 0s are coded as NA's in some cases --> we code all NAs in rows with data as 0s
4) some sites record what veg is present (with "x") --> we replace it by 100%
5) some NAs are coded as character "NA" --> make real NA
6) for some it says "Behöver datajustering"
7) for the remaining cases were the sum is over 110 we set the entries to NA. For sum ≤ 110 we assume rounding errors and keep the number
8) the variable names `perc_veg_0-5` and `perc_veg_5-15` contain "-" which is a problematic character. I rename them to `perc_veg_0_5` and `perc_veg_5_15`

```{r}
RMO <- 
RMO %>% 
  rename("perc_veg_0_5" = `perc_veg_0-5`,
         "perc_veg_5_15" = `perc_veg_5-15`)

#get unique entries to check and fix
RMO_T <- 
  RMO %>% 
  filter(SiteType == "T") %>% #only trnasects should have cover data
  select(SiteID, SiteType, segment, date, starts_with("perc_veg")) %>% 
  distinct %>% 
  group_by(SiteID, SiteType, segment, date) %>% 
  filter(n() > 1) %>% 
  arrange(SiteID, SiteType, segment, date)

RMO_T %>% 
  group_by()

#5) 
RMO_T <- 
  RMO_T %>% 
  mutate(across(starts_with("perc_veg"), ~case_when(.x == "NA" ~ NA_character_,
                                                    TRUE ~ .x)))
#4)
RMO_T <- 
  RMO_T %>% 
  mutate(across(starts_with("perc_veg"), ~case_when(.x %in% c("x", "X") ~ "100",
                                                    TRUE ~ .x)))

#3)
RMO_T <- 
RMO_T %>% 
  rowwise() %>% 
  mutate(across(starts_with("perc_veg"), 
                ~case_when(is.na(.x) & sum(is.na(as.numeric(
                    c_across(starts_with(
                      "perc_veg"))))) <3 ~ "0",
                  TRUE ~ .x))) %>% 
  suppressWarnings()

#6)
RMO_T <- 
  RMO_T %>% 
  mutate(across(starts_with("perc_veg"), ~case_when(.x == "Behöver datajustering" ~ NA_character_,
                                                    TRUE ~ .x)))
#7)
RMO_T <- 
  RMO_T %>% 
  mutate(across(starts_with("perc_veg"), ~case_when(
    sum(
      as.numeric(
        c_across(
          starts_with("perc_veg")))) > 110 ~ NA_character_,
    TRUE ~ .x)))

#convert to numeric 
RMO_T <- 
  RMO_T %>% 
  mutate(across(starts_with("perc_veg"), as.numeric))


#replace in #RMO
RMO <- 
  RMO %>% 
  select(!starts_with("perc_veg")) %>% 
  left_join(RMO_T)

#export sites with veg > 100% for manual check
RMO %>% 
  rowwise() %>% 
  select(SiteID, SiteText, SiteType, segment, date, starts_with("perc_veg")) %>% 
  distinct() %>% 
  mutate(sum_perc = sum(c_across(starts_with("perc_veg")))) %>% 
  filter(sum_perc > 100) %>% 
  filter(sum_perc > 105) %>% 
  arrange(desc(sum_perc)) #%>% 
  write.table("veg_over_100.txt", sep = "\t", row.names = FALSE)
  
RMO %>% 
  ungroup() %>% 
  mutate(rowID = 1:n()) %>% 
  filter(grepl("Djurgården transekter", SiteText)) %>% 
  filter(segment == 2) %>%  
  filter(date == ymd("2016-06-02")) %>% 
  select(rowID, SiteID, segment, date, starts_with("perc_veg")) %>% 
  arrange(rowID)

RMO %>% 
  mutate(rowID = 1:n()) %>% 
  select(!starts_with("perc_veg")) %>% 
  left_join(RMO_T) %>% 
  group_by(rowID) %>% 
  filter(n() > 1)

```


### `per_mill_flowers`

```{r}
#per_mill_flowers --> needs fixing. Includes wrongly formatted numbers, comments and literal "NA" (as text)
RMO$per_mill_flowers[!is.na(RMO$per_mill_flowers)][which(is.na(as.numeric(RMO$per_mill_flowers[!is.na(RMO$per_mill_flowers)])))] %>% table() 
```

1) ,0,05 -> 5%
2) <X -> truncate to X (i.e. strip away "<" or "≤")
3) range (i.e. 15-30) -> replace by mean of range
4) "Behöver datajustering" --> NA
5) "NA" --> NA

```{r}
# 1)
RMO_flow <- 
RMO %>% 
  select(SiteID, SiteType, segment, date, per_mill_flowers) %>% 
  distinct() %>% 
  rowwise() %>% 
  mutate(per_mill_flowers = 
           case_when(per_mill_flowers == ",0,05" ~ "5", #1)
                   grepl("<", per_mill_flowers) ~ gsub("<(\\d)", "\\1", per_mill_flowers), #2)
                   grepl("≤", per_mill_flowers) ~ gsub("≤(\\d)", "\\1", per_mill_flowers), #2)
                   per_mill_flowers == "Behöver datajustering" ~ NA_character_, #4)
                   per_mill_flowers == "NA" ~ NA_character_, #5)
                   grepl("-", per_mill_flowers) ~ as.character(mean(as.numeric(strsplit(per_mill_flowers, "-")[[1]]))), #3)
                   TRUE ~ per_mill_flowers)) %>% 
  suppressWarnings() %>% 
  mutate(per_mill_flowers = as.numeric(per_mill_flowers)) #convert to numeric

#replace fixed data in original dataframe
RMO <- 
  RMO %>% 
  select(-per_mill_flowers) %>% 
  left_join(RMO_flow)
                   
```


### `perc_vadd_tist_klint`

```{r}
#perc_vadd_tist_klint --> needs fixing. Includes wrongly formated NAs and comments
RMO$perc_vadd_tist_klint[!is.na(RMO$perc_vadd_tist_klint)][which(is.na(as.numeric(RMO$perc_vadd_tist_klint[!is.na(RMO$perc_vadd_tist_klint)])))] %>% table() 
```

1) we exclude all non-numeric characters

```{r}
RMO <- 
RMO %>% 
  mutate(perc_vadd_tist_klint = as.numeric(perc_vadd_tist_klint)) #will automatically convert all non-numeric charatcers to NA
```


### `vadd_tist_klint`

TRUE / FALSE /NA. No problems
```{r}
#vadd_tist_klint OK
RMO$vadd_tist_klint %>% unique()
```

### `temp`

numeric or NA, no problem
```{r}
#temp OK
is.numeric(RMO$temp[!is.na(RMO$temp)]) %>% is.na %>% sum
```

### `HavdStatus`

1) some NAs are coded as literal NAs ("NA")

```{r}
RMO$HavdStatus[!is.na(RMO$HavdStatus)] %>% table
```

```{r}
RMO$HavdStatus[!is.na(RMO$HavdStatus) & grepl("NA", RMO$HavdStatus)] <- NA_character_
```


### `Havd_X_kreatur`
```{r}
# HavdNotkreatur
RMO$HavdNotkreatur[!is.na(RMO$HavdNotkreatur)] %>% table
# HavdFar
RMO$HavdFar[!is.na(RMO$HavdFar)] %>% table
# HavdHast
RMO$HavdHast[!is.na(RMO$HavdHast)] %>% table
```

1) some NAs are coded as literal NAs ("NA")
2) the coding differs between Länstyrelsen and SLU: Länstyrelsen record the number of animals whereas SLU records only p/a

However it is not possible to relate the animal number to an area, wherefore the information is not that usefull. 

### `SpeciesText`

1) SpeciesText Liten bastardsvärmare should be changed to Mindre bastardsvärmare, SpeciesID=5
2) SpeciesText Obestämd P. argus should be changed to Ljungblåvinge/Hedblåvinge, SpeciesID=131

3) There are different versions of SpeciesText for the same Species ID (usually first letter capitalized or not). I Keep the  version with the capitalized fist name and change the other names accordingly

4) The `Species ID` variable contains a space. I replace it by `SpeciesID`

```{r}
# 4)
RMO <- 
  RMO %>% 
  rename("SpeciesID" = `Species ID`)

#unique SpeciesID to SpeciesText

Species_ID <- 
RMO %>% 
  mutate(SpeciesID = case_when(SpeciesText == "Obestämd P. argus" ~ "131",
                                  SpeciesText == "Liten bastardsvärmare" ~ "5",
                                  TRUE ~ SpeciesID)) %>% 
  select(SpeciesID, SpeciesText) %>% 
  distinct() %>% 
  #group_by(SpeciesText) %>%      #check if there are NAs in Speciescode for some species names
  #filter(TRUE %in% is.na(SpeciesID))
  group_by(SpeciesID) %>% 
  slice(1) %>% # keep capitalized species name 
  filter(!is.na(SpeciesID)) %>% #does exclude Svävdagsvärmare (no ID)
  arrange(as.numeric(SpeciesID)) %>% 
  mutate(SpeciesText = stringr::str_to_sentence(SpeciesText)) 

#replace SpeciesText with unique and corrected names
RMO <- 
RMO %>% 
  select(-SpeciesText) %>% 
  left_join(Species_ID) 
```

##incorrect observations

From [GitHub](https://github.com/FabianRoger/AoB_Butterflies/issues/3#issue-784563483):

incorrect observations that should be filtered out

```{r}

LL <- list(
  data.frame(t(c("Alkonblåvinge","T",3))),
  data.frame(t(c("Alkonblåvinge","F",3))),
  data.frame(t(c("Asknätfjäril","F",1))),
  data.frame(t(c("Blomvisslare","all",6))),
  data.frame(t(c("Brun gräsfjäril","E",2))),
  data.frame(t(c("Ekspinnare","all",1))),
  data.frame(t(c("Fetörtsblåvinge","all",4))),
  data.frame(t(c("Gullvivefjäril","T",6))),
  data.frame(t(c("Gullvivefjäril","F",6))),
  data.frame(t(c("Friggas pärlemorfjäril","T",2))),
  data.frame(t(c("Friggas pärlemorfjäril","F",2))),
  data.frame(t(c("Grönfläckig vitfjäril","all",30))),
  data.frame(t(c("Krattsnabbvinge","T",4))),
  data.frame(t(c("Kronärtsblåvinge","SLU",1))),
  data.frame(t(c("Ljusgul höfjäril","M",3))),
  data.frame(t(c("Ljusgul höfjäril","T",3))),
  data.frame(t(c("Turkos blåvinge","all",2))),
  data.frame(t(c("Violett guldvinge","all",16))),
  data.frame(t(c("Väpplingblåvinge","T",2))),
  data.frame(t(c("Mnemosynefjäril","E",1)))) 

LL <- 
  bind_rows(LL) %>% 
  mutate()
  
LL_all <- 
LL %>% 
  mutate(X1 = as.character(X1)) %>% 
  filter(X2 == "all") %>% 
  slice(rep(1:n(), each = length(unique(RMO$CountyCode)))) %>% 
  mutate(X2 = rep(unique(RMO$CountyCode), nrow(filter(LL, X2 == "all")))) %>% 
  rbind(c("Kartfjäril", "14", 2)) %>% 
  rbind(c("Midsommarblåvinge", "1", 2)) %>% 
  select(-X3) %>% 
  setNames(., c("SpeciesText", "CountyCode"))
  

Exclude_spec <- 
  LL %>% 
  filter(X2 != "all") %>% 
  rename("county" = X2) %>% 
  left_join(distinct(select(RMO, county, CountyCode))) %>% 
  select(X1, CountyCode) %>% 
  rename("SpeciesText" = X1) %>% 
  rbind(., LL_all)

#exclude a total of 92 observations
RMO <- 
RMO %>% 
anti_join(.,Exclude_spec) 

```

## select variables
```{r}

RMO <- 
RMO %>% select(
  "date", "site code", "SiteText", "SiteID", "SiteType",
  "segment","length",
  "county", "CountyCode", "County_text", "biotope", "biotope_text",
  "perc_veg_0_5", "perc_veg_5_15", "perc_veg_above_15",
  "per_mill_flowers", "perc_vadd_tist_klint", "vadd_tist_klint",
  "HavdStatus", "HavdNotkreatur", "HavdFar", "HavdHast",
  "temp",
  "SpeciesID","SpeciesText",
  "count") %>% 
  arrange(SiteID, segment, date, SpeciesText)
```

## fix spatial Data
```{r}
RMO_spat <- st_as_sf(RMO_spat, coords = c("sit_geort9025gonvlon", "sit_geort9025gonvlat")) 
st_crs(RMO_spat) <- "EPSG:3021"

RMO_spat <- 
RMO_spat %>% 
  rename("SiteID" = sit_uid) %>% 
  select(SiteID, geometry) %>% 
  mutate(SiteID = as.character(SiteID)) %>% 
  left_join(., distinct(select(RMO, SiteID, SiteText, SiteType, county, County_text)))
```

# matching with TUVA objects
```{r}
TUVA <- st_transform(TUVA, st_crs(RMO_spat)) %>% 
  st_make_valid(.)

TUVA <- 
TUVA %>% 
  filter(gsub("(\\w).+", "\\1", Län) %in% unique(RMO_spat$county))

TUVA_RMO <- 
RMO_spat %>% 
  st_buffer(1000) %>% 
  st_intersects(.,TUVA, )

TUVA <- TUVA %>% 
  slice(unique( unlist( TUVA_RMO)))

TUVA_RMO_nearest <- 
RMO_spat %>% 
  st_nearest_feature(.,TUVA)

RMO_spat <- 
TUVA %>% 
  slice(TUVA_RMO_nearest) %>% 
  select(OBJECTID, Markslag) %>% 
  st_set_geometry(NULL) %>% 
  bind_cols(RMO_spat, .)

# check if each SiteText is assigned to a single TUVA object
RMO_spat %>% 
  group_by(SiteText) %>% 
  summarise(unique(OBJECTID)) %>% 
  group_by(SiteText) %>% 
  filter(n() > 1)
  
```



##export
```{r}
save(RMO, RMO_spat, file = "RMO_clean_data.RData")
```

